---
title: "R Notebook"
output: html_notebook
---
## STAT Taxonomy Update - BigQuery 211021

Grabbing STAT analysis file from BigQuery for all SRA analyzed
(see: https://www.ncbi.nlm.nih.gov/sra/docs/aligned-metadata-tables/)


Retrieve STAT taxonomic "Orders", dump the data into CSV.
The `taxid` are then extracted in an ordered list for input to this
scrip

BigQuery: 
`
  SELECT * FROM nih-sra-datastore.sra_tax_analysis_tool.tax_analysis AS tax
  WHERE rank = 'order' AND
      total_count >= 100
  ORDER BY acc
`

SQL calculations:
`
CREATE TABLE sra_stat2 AS
	WITH total_counts AS( 
		SELECT run, SUM(count) AS count_per_run FROM sra_stat
		GROUP BY run
	)
	SELECT sra_stat.*, total_counts.count_per_run,
		ROUND( 100*count/total_counts.count_per_run,2 ) as kmer_perc
		FROM sra_stat
		LEFT OUTER JOIN total_counts on (sra_stat.run = total_counts.run)
`
Update stat table
`
UPDATE sra_stat
SET "group" = sra_stat_group.group
FROM sra_stat_group
WHERE sra_stat.taxid =  sra_stat_group.taxid
`


Make as seperate table
`
CREATE TABLE sra_stat2 AS
	SELECT * 
	FROM sra_stat 
	INNER JOIN add_stat_label ON (sra_stat.row_id = add_stat_label.row_idd)
`

```{r}
# Import NCBI Taxonomy
# install.packages("taxize")
# install.packages("taxizedb")
# Download NCBI Taxonomy
# db_download_ncbi()

# Library
#library("taxize")
library("taxizedb")
library("plyr")

```


```{r}
# Imported ordinal vector of SQL rows for STAT Orders (taxid values)
stat.order <- read.csv(file = "taxid.list", header = FALSE)

# Establish dictionary to be used
stat <- data.frame( taxid    = unique(stat.order),
                    sci_name = as.character(NA),
                    group    = as.character(NA))
  colnames(stat) <- c("taxid", "sci_name", "group")
  
# Initialize 'sci_name'
  stat$sci_name <- taxid2name(stat$taxid)
```  
  
```{r}
# Target space to map NCBI Taxonomy onto for visualizing
target_taxa <- c( "Primate", "Mammal", "Vertebrate",
                  "Arthropod", "Invertebrate",
                  "Fungus",
                  "Plant",
                  "Prokarya",
                  "Virus",
                  "Unclassified")

# target_label | tax_name   | tax_rank
# Primate      - primate    - order
# Mammal       - mammalia   - class
# Vertebrate   - vertebrata - clade
# Arthropod    - arthropoda - phylum
# Invertebrate - metazoa    - kingdom (not above)
# Fungi        - fungi      - kingdom
# Plant        - eukaryota  - superkingdom (not above)
# Archaea      - archaea    - superkingdom
# Bacteria     - bacteria   - superkingdom
# Virus        - viruses    - superkingdom
# Unclassified - NA         - NA
```

```{r}
# Assign Labels Function
assignTaxLabel <- function(taxid) {
  # baseline is superkingdom
  taxn_out <- taxa_at(taxid, rank = "superkingdom", db = "ncbi")[[1]]$name
  
  if (is.null(taxn_out)) {
    return("Unclassified")
    stop()
  }
  
  # kingdom
  kingdom <- taxa_at(taxid, rank = "kingdom", db = "ncbi")[[1]]$name
  
  if (kingdom == "Fungi" || kingdom == "Metazoa") {
    taxn_out <- kingdom
  } else {
    return(taxn_out)
    stop()
  }
  
  # phylum
  phylum <- taxa_at(taxid, rank = "phylum", db = "ncbi")[[1]]$name

  if (phylum == "Arthropoda" || phylum == "Chordata") {
    taxn_out <- phylum
  } else {
    return(taxn_out)
    stop()
  }
  
  # taxclass
  taxclass <- taxa_at(taxid, rank = "class", db = "ncbi")[[1]]$name

  if (taxclass == "Mammalia") {
    taxn_out <- taxclass
  } else if (phylum == "Chordata") {
    # If Chordata is TRUE and Mammalia is False
    # check for Vertebrate
    taxc <- classification(taxid)[[1]]
    
    if ("Vertebrata" %in% taxc$name) {
      taxn_out <- "Vertebrata"
    } else {
      return("Metazoa")
      stop()
    }
  } else {
    return(taxn_out)
    stop()
  }
  
  # taxorder
  taxorder <- taxa_at(taxid, rank = "order", db = "ncbi")[[1]]$name

  if (taxorder == "Primates") {
    taxn_out <- taxorder
  } else {
    return(taxn_out)
  }
}
```

```{r}
# Assign first labels
stat$group <- sapply(stat$taxid, assignTaxLabel)

# Map the original taxid to the labels
stat.assign <- mapvalues(x = stat.order$V1,
                         from = stat$taxid, to = stat$group)
```

```{r}
# Save table
write.table(stat.assign, file = 'label.list', quote = F, row.names = F)

```
